---
title: "Entire home/apt AirBnB Data in NYC"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include = FALSE}
library(flexdashboard)

Packages <- c("tidyverse", "viridis", "dplyr", "readxl", "plotly", "flexdashboard", "p8105.datasets")
invisible(lapply(Packages, library, character.only = TRUE))

data("nyc_airbnb")
set.seed(70)

base_df <- nyc_airbnb %>% 
  janitor::clean_names() %>% 
  rename(
    "rating" = review_scores_location,
    "boro" = neighbourhood_group
  ) %>% 
  mutate(price_per_night = price/minimum_nights,
         rating = rating/2,
         boro = as.factor(boro),
         neighbourhood = as.factor(neighbourhood),
         room_type = as.factor(room_type)
  ) %>% 
  select(boro, neighbourhood, room_type, lat, long, price_per_night, rating) %>% 
  filter(!is.na(rating),
         room_type == "Entire home/apt") %>% 
  sample_n(5000)

```

Column {data-width=650}
-----------------------------------------------------------------------

### Listings with 5-star Ratings in Manhattan by Neighbourhoods

```{r}

ax <- list(
  title = "",
  zeroline = FALSE,
  showline = FALSE,
  showticklabels = FALSE
)

base_df %>% 
  filter(boro == "Manhattan",
         rating == 5) %>% 
  mutate(text_label = str_c("Neighbourhood: ", neighbourhood, "\nPrice/Night: ", price_per_night)) %>%
  plot_ly(x = ~lat, y = ~long, type = "scatter", mode = "markers", 
          color = ~neighbourhood, alpha = 0.5, colors = viridis_pal(option = "D")(31),
          text = ~text_label) %>% 
  layout(xaxis = ax, yaxis = ax)

```

Column {data-width=350}
-----------------------------------------------------------------------

### Neighbourhoods with most <$75.00 home/apt listings

```{r}

base_df %>% 
  filter(price_per_night < 75) %>% 
  group_by(boro, neighbourhood) %>% 
  count(neighbourhood, sort = TRUE) %>% 
  ungroup() %>% 
  top_n(10) %>% 
  droplevels() %>% 
  mutate(neighbourhood = fct_reorder(neighbourhood, n)) %>% 
  plot_ly(x = ~neighbourhood, y = ~n, type = "bar", 
          color = ~boro, colors = viridis_pal(option = "D")(2)) %>% 
  layout(
    xaxis = list(
      title = "Neighbourhoods"),
    yaxis = list(
      title = "Number of < $75.00 listings"
    ))

```

### Distribution of price/night in the 5 Boroughs

```{r}

base_df %>% 
  filter(room_type == "Entire home/apt",
         price_per_night %in% 0:350) %>% 
  mutate(boro = fct_reorder(boro, price_per_night)) %>% 
  plot_ly(y = ~price_per_night, x = ~boro, type = "box", 
          color = ~boro, colors = viridis_pal(option = "D")(5)) %>% 
  layout(
    xaxis = list(
      title = "Boroughs"),
    yaxis = list(
      title = "Price per Night ($)"
    ))

```

