---
title: "how to make websites using rmd"
author: "Kevin S.W."
date: "10/17/2019"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Packages <- c("tidyverse", "viridis", "dplyr", "readxl", "plotly", "flexdashboard", "p8105.datasets")
invisible(lapply(Packages, library, character.only = TRUE))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

# plotly stuff

NOTE: a quick note: we can add code_fold = true to allow hiding/showing codes instead of having to use echo = TRUE, etc

We can use airbnb

```{r}

set.seed(1)

data(nyc_airbnb)
nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(rating = review_scores_location / 2) %>%
  select(
    neighbourhood_group, neighbourhood, rating, price, room_type, lat, long) %>%
  filter(
    !is.na(rating), 
    neighbourhood_group == "Manhattan",
    room_type == "Entire home/apt",
    price %in% 100:500)  %>% 
  sample_n(5000)

```

plotly is more interactive than ggplot. Core structure is similar but the way you use plotly is different...

##### plotly scatterplot

```{r}

nyc_airbnb %>%
  mutate(text_label = str_c("Price: $", price, '\nRating: ', rating)) %>%  # concatenate stuff into the texts
  plot_ly(
    x = ~lat, y = ~long, type = "scatter", mode = "markers",
    color = ~price, text = ~text_label, alpha = 0.5) # "~" indicates the "variable" you need for plotly

```

##### plotly boxplots...

```{r}

common_neighborhoods =
  nyc_airbnb %>% 
  count(neighbourhood, sort = TRUE) %>% 
  top_n(8) %>% 
  select(neighbourhood)

nyc_airbnb %>% 
  mutate(neighbourhood = fct_reorder(neighbourhood, price)) %>% 
  plot_ly(y = ~price, color = ~neighbourhood, type = "box",
          colors = "Set2")


inner_join(nyc_airbnb, common_neighborhoods, by = "neighbourhood") %>% 
  mutate(neighbourhood = fct_reorder(neighbourhood, price)) %>% 
  plot_ly(y = ~price, color = ~neighbourhood, type = "box",
          colors = "Set2")

```

##### barcharts...

```{r}

nyc_airbnb %>% 
  count(neighbourhood) %>% 
  mutate(neighbourhood = fct_reorder(neighbourhood, n)) %>% 
  plot_ly(x = ~neighbourhood, y = ~n, color = ~neighbourhood, type = "bar")

```

##### scatterplots....

```{r}

scatter_ggplot = 
  nyc_airbnb %>%
  ggplot(aes(x = lat, y = long, color = price)) +
  geom_point(alpha = 0.25) +
  scale_color_viridis() +
  coord_cartesian() +
  theme_classic()

ggplotly(scatter_ggplot)

```

##### you could also "modify" ggplots to make it more interactive

Though since plot_ly is built with interactivity in mind and ggplot isn't, ggplot will be less "beautiful". 

```{r}

box_ggplot = 
  inner_join(nyc_airbnb, common_neighborhoods,
             by = "neighbourhood") %>% 
  mutate(neighbourhood = fct_reorder(neighbourhood, price)) %>% 
  ggplot(aes(x = neighbourhood, y = price, fill = neighbourhood)) +
  geom_boxplot() +
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplotly(box_ggplot)

```

# Using flexdashboard

Clearly you can embed interactive graphics in HTML files produced by R Markdown; this is a handy time to introduce dashboards. In short, dashboards are a collection of related graphics (or tables, or other outputs) that are displayed in a structured way that’s easy to navigate.

You can create dashboards using the flexdashboard package by specifying flex_dashboard as the output format in your R Markdown YAML. There are a variety of layout options, but we’ll focus on a pretty simple structure produced by the template below. This is the default dashboard template in R Studio – if you have flexdashboard installed, you can use File > New File > R Markdown > From Template to create a new .Rmd file with the structure below.

Now, when making dashboard, you save it just outside of the project folder first. Reason is because, when you knit a dashboard inside a "web", there's a particular way to do it.

***NOTE***

Reason being, when you knit inside the project, by default, knitr will just "strip" the "dashboard" components; the website’s _site.yml file conflicts with the dashboard’s YAML header regarding the output format – and the website’s _site.yml “wins”. To address this issue, instead of knitting you can use this command to knit the dashboard.

`rmarkdown::render("dashboard_template.Rmd", output_format = "flexdashboard::flex_dashboard")`

This will create dashboard_template.html but not open it in RStudio’s Viewer pane; you can open the file in a browser instead. Alternatively, using RStudio’s Build pane to Build Website will produce the same results. To illustrate, we’ll put the dashboard we just created on a website for this topic.

All of this YAML business is only an issue for dashboards embedded in websites; a standalone dashboard (in a non-website GH repo or R Project) can be knit using the same process as other .Rmd file
