---
title: "Plotting with Plotly"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

Packages <- c("tidyverse", "viridis", "dplyr", "readxl", "plotly", "flexdashboard", "p8105.datasets")
invisible(lapply(Packages, library, character.only = TRUE))

```

<br></br><br></br>

# Preparing Datasets
We're going to use the NYC AirBnB dataset in `p8105.datasets` from our P8105 class. More information about the dataset is [here](https://p8105.com/dataset_airbnb.html). The plots can be seen [here](dashboard-hw.html) (or click on "Stuff I Learned" and click on "Dashboard"). This page will describe the code behind the graphs.

```{r data_cleaning}

data("nyc_airbnb")
set.seed(70)                    # fix randomization

base_df <- nyc_airbnb %>% 
  janitor::clean_names() %>% 
  rename(
    "rating" = review_scores_location,                 # rename reviews to an easier name
    "boro" = neighbourhood_group                       # rename to a more familiar term
  ) %>% 
  mutate(price_per_night = price/minimum_nights,       # add price/night column
         rating = rating/2,                            # turn "10-star" rating to "5-star"
         boro = as.factor(boro),                       # turns variables to factors
         neighbourhood = as.factor(neighbourhood),
         room_type = as.factor(room_type)
  ) %>% 
  select(boro, neighbourhood, room_type, lat, long,    # select relevant variables
         price_per_night, rating) %>%  
  filter(!is.na(rating),                               # filter out NA
         room_type == "Entire home/apt") %>%           # only data from entire home/apt
  sample_n(5000)                                       # obtain 5000 samples randomly

```

We filter out unnecessary variables and retains only the generally useful information: boroughs, neighbourhoods, room type, latitude/longitude, price per night, and 5-star rating. Since our original dataframe contains a vast amount of data, we took a smaller sample of `r nrow(base_df)`. 
<br></br>

# Getting plot for distribution of nightly price

```{r, night_price_box, eval = FALSE}

base_df %>% 
  filter(price_per_night %in% 0:350) %>%                              # narrows price range to $0-$350 / night
  mutate(boro = fct_reorder(boro, desc(price_per_night))) %>%         # re-orders boroughs based on price
  plot_ly(y = ~price_per_night, x = ~boro, type = "box",              # make boxplot
          color = ~boro, colors = viridis_pal(option = "D")(5)) %>%   # tells viridis to get 5 colors
  layout(
    xaxis = list(
      title = "Boroughs"),                                            # add axis labels
    yaxis = list(
      title = "Price per Night ($)"
    ))

```

We then create a simple distribution boxplot that shows how the nightly price for entire home/apt differs by boroughs. Though be mindful that we only take a sample of `r nrow(base_df)`.
<br></br>

# Making bar charts for top 10 neighbourhoods with nightly prices below $75 

```{r top10bar, eval = FALSE}

base_df %>% 
  filter(price_per_night < 75) %>%                                    # filter for prices < $75.00
  group_by(boro, neighbourhood) %>%                                   # group by borough and neighbourhood
  count(neighbourhood, sort = TRUE) %>%                               # count the listings in each neighbourhood
  ungroup() %>%                         
  top_n(10) %>%                                                       # obtain the top 10 with most listings 
  droplevels() %>%                                                    # remove unused levels
  mutate(neighbourhood = fct_reorder(neighbourhood, n)) %>%           # reorder neighbourhood by count
  plot_ly(x = ~neighbourhood, y = ~n, type = "bar",                   # plot the bargraph
          color = ~boro, colors = viridis_pal(option = "D")(2)) %>% 
  layout(
    xaxis = list(                                                     # axis labels
      title = "Neighbourhoods"),
    yaxis = list(
      title = "Number of < $75.00 listings"
    ))

```

This barchart shows the top 10 neighbourhoods with nightly prices of < $75.00.
<br></br>

# Scatterplot of listings in Manhattan with ratings 4 and higher

```{r scatter5star, eval = FALSE}

ax <- list(                     # setting to remove grid, labels
  title = "",
  zeroline = FALSE,
  showline = FALSE,
  showticklabels = FALSE,
  showgrid = FALSE
)

base_df %>% 
  filter(boro == "Manhattan",                                        # filter by Manhattan and 5-star rating
         rating == 5) %>% 
  mutate(
    text_label = str_c("Neighbourhood: ", neighbourhood,             # assigns hovering-label
                       "\nPrice/Night: ", price_per_night)) %>%  
  plot_ly(x = ~lat, y = ~long, type = "scatter", mode = "markers",   # plots the scatter plot
          color = ~neighbourhood, alpha = 0.5, colors = viridis_pal(option = "D")(31),
          text = ~text_label) %>% 
  layout(xaxis = ax, yaxis = ax)

```

This scatterplot code uses the longitude/latitude points to make a "map" of the 5-stars in Manhattan. Hovering over it shows the location and price/night of that location.
