---
title: "Plotting with Plotly"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

Packages <- c("tidyverse", "viridis", "dplyr", "readxl", "plotly", "flexdashboard", "p8105.datasets")
invisible(lapply(Packages, library, character.only = TRUE))

```

<br></br><br></br>

# Preparing Datasets
We're going to use the NYC NOAA dataset in `p8105.datasets` from our P8105 class. More information about the dataset is [here](https://p8105.com/dataset_noaa.html). The plots can be seen [here](dashboard-hw.html) (or click on "Stuff I Learned" and click on "Dashboard"). This page will describe the code behind the graphs.

```{r data_cleaning}

data("ny_noaa")

base_df <- ny_noaa %>% 
  janitor::clean_names() %>% 
  drop_na() %>%                                                     # drop any NA in any column
  filter(id %in% c("USW00094728", "USC00300961", "USC00306164", "USC00307799", "USW00014735")) %>% 
  mutate(location = case_when(id == "USW00094728" ~ "Manhattan",
                              id == "USC00300961" ~ "Bronx",
                              id == "USC00306164" ~ "Ogdensburg",
                              id == "USC00307799" ~ "Slide Mountain",
                              id == "USW00014735" ~ "Albany"),
         tmin = as.numeric(tmin)/10,                                # convert from tenths to regular C
         tmax = as.numeric(tmax)/10,
         year = lubridate::year(date),                              # turn to specified date format
         month = lubridate::month(date),
         date = lubridate::day(date)) %>% 
  select(location, year, month, date, everything(), -id, -snwd)

```

We filter out unnecessary variables and retains only the information we want: year/month/date, precipitation, snowfall, and max/min temp. We will be comparing these information between locations in New York State; Manhattan, Bronx, Ogdensburg, Slide Mountain, and Albany.
<br></br>

# Getting plot for annual snowfall between `r min(pull(base_df, year))` - `r max(pull(base_df, year))`

```{r annual_snow, eval = FALSE}

base_df %>% 
  select(location, year, month, snow) %>%
  group_by(location, year) %>% 
  summarize(yearly_snow = sum(snow)) %>% 
  plot_ly(y = ~yearly_snow, x = ~year, type = "scatter", mode = "lines",     # make line graph
          color = ~location, colors = viridis_pal(option = "D")(5)) %>%      # tells viridis to get 5 colors %>% 
  layout(
    xaxis = list(
      title = "Year",
      autotick = FALSE, tick0 = 1980, dtick = 2, tickangle = 45),            # manually assigns ticks/tick-gaps
    yaxis = list(
      title = "Annual Snowfall (mm)"
    ))

```

This particular code creates a simple line graph with colors based on the locations that shows how the annual snowfall fluctuates over the years.
<br></br>

# Manhattan Monthly Temperature (C) changes in 2010

```{r monthly_temp, eval = FALSE}

base_df %>% 
  select(location, year, month, tmin, tmax) %>%
  filter(location == "Manhattan",
         year == 2010) %>%
  pivot_longer(tmax:tmin,
               names_to = "obs",
               values_to = "temp") %>% 
  plot_ly(y = ~temp, x = ~month, type = "scatter", mode = "markers",     
          color = ~obs, colors = viridis_pal(option = "D")(10), opacity = 0.4) %>%
  layout(
    xaxis = list(
      title = "Months in 2010",
      autotick = FALSE, tick0 = 1, dtick = 1),
    yaxis = list(
      title = "Temperature (C)"
    ))

```

Focusing on Manhattan (Central Park) itself, we can also check the monthly changes in minimum and maximum temperature each year using a scatterplot. In this case, we particularly focus on 2010. 
<br></br>

# Bronx's Total Precipitation and Average Min Temperature in 2010

```{r precip_tmin, eval = FALSE}

base_df %>% 
  select(location, year, month, prcp, tmin) %>%
  filter(location == "Bronx",
         year == 2010) %>% 
  group_by(month) %>% 
  summarize(total_prcp = sum(prcp),
            avg_tmin = mean(tmin)) %>% 
  plot_ly() %>% 
  add_trace(x = ~month, y = ~total_prcp, type = "bar", name = "Precipitation (mm)",  # add bar graph
            hoverinfo = "text",                                                      # add hoverinfo
            text = ~paste(total_prcp, " mm")) %>%                                    # hoverinfo content
  add_trace(x = ~month, y = ~avg_tmin, type = "scatter", mode = "line", name = "Min Temp (C)", # add line graph
            yaxis = "y2",                                                                      # specify 2nd y-axis
            hoverinfo = "text",
            text = ~paste(avg_tmin, " C")) %>% 
  layout(
    yaxis = list(
      side = 'left', 
      title = 'Precipitation (mm)', 
      showgrid = FALSE, zeroline = FALSE),
    yaxis2 = list(
      side = 'right', overlaying = "y", 
      title = 'Min Temp (C)', 
      showgrid = FALSE, zeroline = FALSE),
    xaxis = list(
      title = "Months in 2010",
      autotick = FALSE, tick0 = 1, dtick = 1, tickangle = 0)
  )

```

This code combines a bar plot (for total precipitation) and line plot (for average min temp) in order to see possible correlation with temperature and "rainfall" in Bronx, 2010. 
